---
title: "B_Plotting individual UDs"
format: html
editor: source
editor_options: 
  chunk_output_type: console
---

[ARG: Devin developed the code for Group UDs prior to the code for individual UD, thus the latter actually depends on steps in the former. We don't want this...we want to be able to calculate individual UDs (hence why I have made this step 4 here) and the user can calculate group UDs if they want (using step 5).]{.underline}

### Read in saved data and load packages
If you previously stopped after step 2 and want to restart, you have to read in the saved data and load all the packages.

```{r}
#| output: false
load("processed_data/ud_df.RData")
packages <- read.csv("packages.csv")$packages
invisible(lapply(packages, library, character.only = TRUE))
```

Read in the metadata again. If you want you could just save from the beginning.

```{r}
#| output: false
meta_data <- read_csv("raw_data/Metadata.csv") %>% 
  clean_names() %>% rename(deploy_id = ptt_unique_id)

meta_data <- meta_data %>% filter(deploy_id %in% ud_df$deploy_id)

ud_data <- select(ud_df, deploy_id, data, ess, area95, area50, ud_full, ud_95, ud_50)
ud_data <- left_join(ud_df, meta_data)

```

### Capture location spatial points

[ARG: The code in the first chunk below is only in the UD plots for all individual turtles being processed at the same, it isn't in the later code where we could calculate UD for each individual turtle, or the Group UD code.]{.underline}

```{r}
capt_loc <- meta_data %>% select(deploy_id, capture_longitude, capture_latitude) %>% 
  st_as_sf(coords = c("capture_longitude","capture_latitude"), crs=4326)
capt_loc <- st_transform(capt_loc, st_crs(land))
```


### Create output folder

Adjust names of files as necessary

```{r}
if(!dir.exists("plots/individual_uds/joint")){dir.create("plots/individual_uds/joint")}

```

### Get Bathymetry maps for plotting 

```{r}
library(rosm)
locs_bbox <- st_expand(st_bbox(land), 0.2) %>% st_as_sfc()
esri_ocean <- source_from_url_format(
  url_format = paste0('https://services.arcgisonline.com/arcgis/rest/services/',
                      'Ocean/World_Ocean_Base/MapServer/tile/${z}/${y}/${x}.jpeg'),
  attribution = "")

esri_tiles <- osm.raster(sf::as_Spatial(locs_bbox),
                         type = esri_ocean, zoomin = 2,
                         progress = "none", crop = FALSE)
esri_tiles <- terra::rast(esri_tiles)
terra::crs(esri_tiles) <- "EPSG:6637"
```

### Indicate tag(s) to be processed for UDs

Can calculate UDs for all tags in a data set or for individuals tags. If you want to plot a specific PTT, just use that number for the `PTT <- ...` in the loop below and don't run the whole loop just the code within the loop.   

[To calculate the UDs of all tags in a data set AT ONCE, use the code in this chunk (skip to next chunk for individual tags):]{.underline}

```{r}
for(i in 1:nrow(ud_data)){
  
  PTT <- ud_data$deploy_id[[i]]
  ud_single_df <- ud_data %>% filter(deploy_id==PTT)
  ud <- readRDS(paste0("ud_files/ud_full/",PTT,".rds"))
  ud_cont <- cu_ud_contour(ud, c(0.95,0.5))
  ud_cont_poly <- ud_cont %>% st_cast("POLYGON")
  ## Capture locs
  capt_pts <- capt_loc %>% filter(deploy_id==PTT)
  pts <- st_as_sf(ud_single_df$data[[1]][,"geometry"])
  # Plots
  bbb <- st_buffer(ud_cont[1,], 1500) |> st_bbox() |> st_as_sfc()
  
  ###
  ### See http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf
  ### if you want to change homerange colors and 
  ### https://colorbrewer2.org/ for the PTT color palette
  ###
  ppp <- ggplot() +
    annotation_spatial(esri_tiles) +
    layer_spatial(bbb, fill=NA, size=0, color=NA) +
    annotation_spatial(pts, size=0.25, alpha=0.5, color="black") + 
    layer_spatial(ud_cont_poly[1,], color="royalblue", lwd=0.25, fill="royalblue", alpha=0.3) +      # 95% quantile polygon
    layer_spatial(ud_cont_poly[2,], color="firebrick2", lwd=0.25, fill="firebrick2", alpha=0.3) +    # 50% quantile polygon
    annotation_spatial(land, fill = "gray50", size = 0) + # land
    layer_spatial(capt_pts, size=2, alpha=1, color="green", shape=3) +
    theme_bw() + theme(panel.grid.major = element_blank(), 
                       panel.grid.minor = element_blank(),
                       axis.ticks=element_blank(), axis.text=element_blank()) +
    annotation_scale()
  
  ggsave(ppp, file=paste0("plots/individual_uds/joint/", PTT,"_contour.png"), width=6.5, height=4)
}
```

