---
title: "D_Land maps"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

[ARG. The Island Plots file Devin created allows one to produce maps of the Marianas (each island individually or the entire chain), which can be used when making maps (e.g., UD maps can be inset on these larger, island-wide maps for creating figures).]{.underline}

[We need to edit this code so that we can easily create maps for anywhere (for use in making figures)]{.underline}

### Read in saved data and load packages
If you previously stopped after step 3 and want to restart, you have to read in the saved data and load all the packages.

```{r}
load("processed_data/predict_samples.RData")
packages <- read.csv("packages.csv")$packages
invisible(lapply(packages, library, character.only = TRUE))
```

### Load data if not already in the workspace

Adjust the names of the files in below chunk based on what was created/saved in previous steps

```{r}
load("processed_data/predict_samples_NavyMarianas.RData")
```

### Create land polygon

```{r}
locs_bbox <- st_expand(st_bbox(land), 0.2) %>% st_as_sfc()
esri_ocean <- source_from_url_format(
  url_format = paste0('https://services.arcgisonline.com/arcgis/rest/services/',
                      'Ocean/World_Ocean_Base/MapServer/tile/${z}/${y}/${x}.jpeg'),
  attribution = "")

esri_tiles <- osm.raster(sf::as_Spatial(locs_bbox),
                         type = esri_ocean, zoomin = 2,
                         progress = "none", crop = FALSE)
esri_tiles <- terra::rast(esri_tiles)
terra::crs(esri_tiles) <- "EPSG:6637"

land <- st_cast(land, "POLYGON")
land$FID <- 1:nrow(land)
```

### Hover over map to see polygon numbers

```{r}
mapview(land)
```

### Create map for a specific area

Can adjust by changing the polygon numbers in the first line

This code creates a map for Rota:

```{r}
rota <- st_geometry(land[43,]) %>% st_buffer(20000) %>% st_intersection(land)
bbb <- st_bbox(rota) %>% st_as_sfc() %>% st_buffer(5000) # add a little buffer 5km
ppp <- ggplot() +
  annotation_spatial(esri_tiles) +
  layer_spatial(rota, fill = "gray50", size = 0) +
  layer_spatial(bbb, fill=NA, size=0, color=NA) +
  annotation_scale() +
  theme_bw() 
ggsave(ppp, file=paste0("plots/rota.png"), width=6, height=6.5, dpi="retina")

```

This code is for Tinian and Saipan:

```{r}
## Tinian/Saipan
tinian_saipan <- st_geometry(land[c(48,54),])  %>% st_buffer(20000) %>% st_intersection(land)
bbb <- st_bbox(tinian_saipan) %>% st_as_sfc() %>% st_buffer(5000) # add a little buffer 5km
ppp <- ggplot() +
  annotation_spatial(esri_tiles) +
  layer_spatial(tinian_saipan, fill = "gray50", size = 0) +
  layer_spatial(bbb, fill=NA, size=0, color=NA) +
  annotation_scale() +
  theme_bw() 
ggsave(ppp, file=paste0("plots/tinian_saipan.png"), width=6, height=6.5, dpi="retina")

```

This code is for Guam:

```{r}
## Guam
guam <- st_geometry(land[41,]) %>% st_buffer(20000) %>% st_intersection(land)
bbb <- st_bbox(guam) %>% st_as_sfc() %>% st_buffer(5000) # add a little buffer 5km
ppp <- ggplot() +
  annotation_spatial(esri_tiles) +
  layer_spatial(guam, fill = "gray50", size = 0) +
  layer_spatial(bbb, fill=NA, size=0, color=NA) +
  annotation_scale() +
  theme_bw() 
ggsave(ppp, file=paste0("plots/tinian_saipan.png"), width=6, height=6.5, dpi="retina")

```

